---
title: "Logistic Regression"
author: "Clinton Moshe"
date: "10/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the Data and packages

```{r data}
library(tidyverse)
library(caret) # tuning, splitting the data etc

AdmData <- read_csv('data/Admission_Predict_Ver1.1.csv')[-1]
str(AdmData)
# we will convert them appropriately later on
summary(AdmData)
```

## Data Preparation

```{r data prep}
# converting them into the correct data types
# i have renamed(AdmData --> AdmData_1) to preserve the original data
AdmData_1 <- AdmData %>%
  mutate(
    `University Rating` = as_factor(`University Rating`),
    `Research` = as_factor(`Research`),
    `SOP` = as_factor(`SOP`)
  )
str(AdmData_1)
summary(AdmData_1)

# checking NA/MISSING DATA
colSums(is.na(AdmData_1)) # no NA values
AdmData_1$`Chance of Admit` # to help me choose a representative threshold

# split the data
AdmData_1$`Chance of Admit` <- ifelse(AdmData_1$`Chance of Admit` >0.7,1,0)

# index to split
index <- createDataPartition(AdmData_1$`Chance of Admit`,list = FALSE, p=0.8)
index = as_vector(index)
train <- AdmData_1[index,]
test <- AdmData_1[-index,]
# more training data for better examples for the model to learn from
```

## Exploratory Data Analaysis

```{r EDA}
# i willnot do EDA(exploratory data analysis but you can do it here)
# use ggplot2, it is already loaded.

```

## Modelling

```{r modelling}
# adding the train data into the search path
attach(train)
model1 <- glm(`Chance of Admit`~., 
              data = AdmData_1, 
              family = poisson(link = "log"))
summary(model1)
# make many models for the sake of choosing the better model
# i am going to use model1 to predict for this video
```

## making predictions

```{r predictions}
options(scipen=999) # disable scientifc notation
pred1 <- predict(model1, newdata = test, type = "response")
pred1 <- ifelse(pred1 >=0.7,1,0)

# making comparison
df <- data.frame(pred1, test$`Chance of Admit`)
df
tb <- table(test$`Chance of Admit`, pred1)
confusionMatrix(tb)
```
# you can create another model and try to tune using caret package..
# also try to normalize, standardize the data to see if it will have any
# impact on the performance of the model.
# i will leave the link on the description below
# THANKS GUYS AND REMEMBER TO SUBSCRIBE AND SHARE.
